The Near and Mid-infrared photometric properties of known redshift $z\geq5$ Quasars

### 2020-02-15 tex/MNRAS_submission/rv1/VHzQ_MNRAS_rv1.tex
Updated a few numbers in Sect 3.3,3.4 and Conclusions

The set of VH$z$Q detected in WISE W3 and W4 contains 51 objects that
are detected in the broad W3 filter and 10 objects that are formally
detected in W4 ({\tt w4snr}$\geq$3.0). Three of the W3/4 objects are
presented in Figure~\ref{fig:VHzQ_W4s}. UHS~J0439+1634
(Fig.~\ref{fig:VHzQ_W4s} {\it (b)}) with W4=7.165$\pm$0.12 was
discovered by \citet{Fan2019} and is a strongly lensed quasar at
$z=$6.51. This high luminosity is mostly not intrinsic, but is boosted
by an intervening redshift $z\sim$0.7 galaxy. Other high-redshift
quasars that are bright W3/4 objects e.g. SDWISE~J0306+1853
(Fig.~\ref{fig:VHzQ_W4s} {\it(c)}) may well also so be lensed
\citep[e.g.,][]{Glikman2018lens, Fan2019, Pacucci2019}, but
high-resolution follow-up is needed to confirm. SDSS~J0100+2802
(redshift $z=6.33$; Fig.~\ref{fig:VHzQ_W4s} {\it(d)}) does not have a
formal W4 detection ({\tt w4snr}=2.4), but as reported by
\citet{Wu2015}, has a detection in the 2MASS $J$, $H$ and $K_{\rm s}$
bands and is also the bluest object in both (W1-W3) and (W2-W3).

@@

Figure~\ref{fig:MIR_LC_epochs} gives the number of NEOWISE-R epochs and detections for each VH$z$Q. 
There are 60 quasars with over 100 NEOWISE-R epochs, nine with over 200 epochs and four with over 300.
Figure~\ref{fig:MIR_LC_3egs} presents three examples of the brighter MIR objects with well-sampled 
MIR lightcurves and associated colour changes; 
SDSS J0100+2802 (Fig.~\ref{fig:MIR_LC_3egs} top), 
SDWISE  J1752+5036    (middle), 
SDSS J1621+5155 (bottom). 
For the light curves, we show weighted means for every 6 month NEOWISE-R scan.
Due to the WISE scanning pattern, objects close to the North Ecliptic Pole 
are well sampled, and indeed this is the area for further investigation 
time-domain investigation with upcoming large infrared space telescopes
\citep{Jansen2018, WhiteC2019AAS}.

@@

\begin{itemize}
  \item SDSS and Pan-STARRS1 together identified over half of the VH$z$Q sample;
  \item There remains a quasar ``redshift desert'' at $z\approx5.3-5.7$,  though
  efforts are being made to address this \citep[e.g.,][]{YangJ2018a};
  \item 96.9\% of the VH$z$Q sample is detected in one or more NIR
    ($ZYJHK/K_{s}$) band;
  \item The 15 objects that are not detected in the NIR are due to
    lack of coverage rather than lack of depth;
  \item 362 (78.2\%) VH$z$Qs are detected by WISE, e.g. in the deeper
    unWISE W1 catalog.
  \item All of the $z\geq7$ quasars are detected in both unWISE W1 and
    W2.
  \item 28 of the quasars had enough NIR measurements and sufficient
    NIR measurements and signal-to-noise ratio to look for variability. Weak
    variability was detected in multiple bands of SDSS J0959+0227, and
    very marginally in the Y-band of MMT J0215-0529. Only one quasar, SDSS
    J0349+0034 had significant differences between their WFCAM and VISTA
    magnitudes in one band, indicating variability.
 \end{itemize}

>>>>

The set of VH$z$Q detected in WISE W3 and W4 contains 59 objects that
are detected in the broad W3 filter and 10 objects that are formally
detected in W4 ({\tt w4snr}$\geq$3.0). Three of the W3/4 objects are
presented in Figure~\ref{fig:VHzQ_W4s}. UHS~J0439+1634
(Fig.~\ref{fig:VHzQ_W4s} {\it (b)}) with W4=7.165$\pm$0.12 was
discovered by \citet{Fan2019} and is a strongly lensed quasar at
$z=$6.51. This high luminosity is mostly not intrinsic, but is boosted
by an intervening redshift $z\sim$0.7 galaxy. Other high-redshift
quasars that are bright W3/4 objects e.g. SDWISE~J0306+1853
(Fig.~\ref{fig:VHzQ_W4s} {\it(c)}) may well also so be lensed
\citep[e.g.,][]{Glikman2018lens, Fan2019, Pacucci2019}, but
high-resolution follow-up is needed to confirm. SDSS~J0100+2802
(redshift $z=6.33$; Fig.~\ref{fig:VHzQ_W4s} {\it(d)}) does not have a
formal W4 detection ({\tt w4snr}=2.4), but as reported by
\citet{Wu2015}, has a detection in the 2MASS $J$, $H$ and $K_{\rm s}$
bands and is also the bluest object in both (W1-W3) and (W2-W3).

@@

Figure~\ref{fig:MIR_LC_epochs} gives the number of NEOWISE-R epochs and detections for each VH$z$Q. 
There are 74 quasars with over 100 NEOWISE-R epochs, nine with over 200 epochs
and four with over 300.
Figure~\ref{fig:MIR_LC_3egs} presents three examples of the brighter MIR objects with well-sampled 
MIR lightcurves and associated colour changes; 
SDSS J0100+2802 (Fig.~\ref{fig:MIR_LC_3egs} top), 
SDWISE  J1752+5036    (middle), 
SDSS J1621+5155 (bottom). 
For the light curves, we show weighted means for every 6 month NEOWISE-R scan.
Due to the WISE scanning pattern, objects close to the North Ecliptic Pole 
are well sampled, and indeed this is the area for further investigation 
time-domain investigation with upcoming large infrared space telescopes
\citep{Jansen2018, WhiteC2019AAS}.

@@

\begin{itemize}
  \item SDSS and Pan-STARRS1 together identified over half of the VH$z$Q sample;
  \item There remains a quasar ``redshift desert'' at $z\approx5.3-5.7$,  though
  efforts are being made to address this \citep[e.g.,][]{YangJ2018a};
  \item 96.9\% of the VH$z$Q sample is detected in one or more NIR
    ($ZYJHK/K_{s}$) band;
  \item The 15 objects that are not detected in the NIR are due to
    lack of coverage rather than lack of depth;
  \item 389 (79.7\%) VH$z$Qs are detected by WISE, e.g. in the deeper
    unWISE W1 catalog.
  \item All of the $z\geq7$ quasars are detected in both unWISE W1 and
    W2.
  \item 28 of the quasars had enough NIR measurements and sufficient
    NIR measurements and signal-to-noise ratio to look for variability. Weak
    variability was detected in multiple bands of SDSS J0959+0227, and
    very marginally in the Y-band of MMT J0215-0529. Only one quasar, SDSS
    J0349+0034 had significant differences between their WFCAM and VISTA
    magnitudes in one band, indicating variability.
 \end{itemize}
